---
title: "download_data"
author: "steppe"
date: "4/8/2022"
output: html_document
---


You will need to download a vector file which contains the BLM office boundaries.
```{r Load packages, message = F, warning = F}
library(tidyverse)
library(here)
library(FedData)
library(sf)
library(raster) # needed for gdistance
library(terra)
library(whitebox)
library(gdistance)
source(here::here('scripts/functions.R'))
```

```{r Read in and Subset BLM field offices}
FO <- st_read(paste0(here(), "/data/raw/blm_co_admu_fo_shp/BLM_CO_ADMU_FO_20201001.shp"),
              quiet = T)

FO <- FO %>% 
  filter(ADMU_NAME == 'UNCOMPAHGRE FIELD OFFICE') %>% 
  st_transform(4269) %>% 
  as('Spatial')
```


```{r Read in AIM points}

```


### Slope Calculations 

This code serves to download 10m resolution DEM. 
```{r download DEM, eval = F}
#dir.create(paste0(here(), '/data/raw/DEM'))
links <- read.csv(
  paste0(here(), '/data/raw/', 'DEM_file_paths.csv'), header = F)[,c(1,15)]
names(links) <- c('Tile', 'URL')

links <- links %>% 
  mutate(Tile = str_extract_all(Tile, 'n\\d{2}w\\d{3}')) %>% 
  mutate(Tile = paste0(Tile, '_10m_DEM.tif'))

destfile <- paste0(here(), '/data/raw/DEM/', links$Tile)
download.file(links$URL[1], destfile[1])

for (i in 1:length(destfile)){
  
  download.file(links$URL[i], destfile[i])
  Sys.sleep(5)
  message("File downloaded and saved to location: " destfile[i])
}
rm(links, destfile, i)
```

Here we crop the extent of the rasters, and mosaic them together to form a a raster for easier usage.
```{r Load and Mosaic rasters into Spatstat object, eval = F}
raster_files <- paste0(here(), '/data/raw/DEM/', 
                       list.files(paste0(here(), '/data/raw/DEM/')))

raster_files = lapply(raster_files, rast)
rsrc <- sprc(raster_files)
m <- mosaic(rsrc)

FO <- FO %>% 
  st_as_sf() %>% 
  st_transform(32613) %>% 
  st_buffer(80) %>%  # we are just going to buffer a little bit to make sure our edges are nice
  st_transform(crs(m)) %>% 
  vect()

dem <- crop(m, FO)
dem <- terra::mask(dem, FO)

writeRaster(dem, paste0(here(), '/data/raw/UFO_dem_10m.tif'))
rm(raster_files, rsrc, m)
```

We will finally create a raster with modeled slope values. 
```{r Create Slope Raster, eval = F}
dem <- rast(paste0(here(), '/data/raw/UFO_dem_10m.tif'))

terrain(dem, v = 'slope', neighbors = 8, unit = 'degrees', 
               filename = paste0(here(), '/data/processed/UFO_slope_deg.tif')
               )
terrain(dem, v = 'aspect', neighbors = 8, unit = 'degrees', 
               filename = paste0(here(), '/data/processed/UFO_aspect.tif')
               )
terrain(dem, v = 'TRI', neighbors = 8, unit = 'degrees', 
               filename = paste0(here(), '/data/processed/UFO_topo_roughness.tif')
               )

```

Points with greater than 26.57 degrees (slopes of 50*) are office rejections.
```{r Read in Slope degree raster, eval = F}
degrees<- rast(paste0(here(), '/data/processed/UFO_slope.tif'))
plot(degrees)
#aspect<- rast(paste0(here(), '/data/processed/UFO_aspect.tif'))
#plot(aspect)
#tri <- rast(paste0(here(), '/data/processed/UFO_topo_roughness.tif'))
#plot(tri)

rm(aspect, tri)
```


```{r Create Geomorphons and Hydrologic Basins of Field Office, eval = F}
data_proc <- paste0(here(),"/data/processed/")

Sys.time()
whitebox::wbt_feature_preserving_smoothing(paste0(here(),'/data/raw/UFO_dem_10m.tif'), 
                           paste0(data_proc, "UFO_dem_10_smooth.tif"), filter=9)

wbt_breach_depressions(paste0(data_proc, "UFO_dem_10_smooth.tif"), 
                   paste0(data_proc, "UFO_dem_10_sm_breached.tif"))

wbt_geomorphons(paste0(data_proc, "UFO_dem_10_sm_breached.tif"), 
                output = paste0(data_proc, "UFO_geomorphons_10m.tif"), 
                search = 50, forms = T)
Sys.time()


geomor <- raster::raster(paste0(data_proc, "UFO_geomorphons_10m.tif"))
plot(geomor)

rm(data_proc)
```


As a future endeavor we can use AIM data to test our slope value raster. Note this step is only possible if you have had previous AIM work carried out in the area. 

```{r Logistic Regression Determining Accuracy of classification, eval = F}

```


### Pull in and Process Road Data

```{r}
john_tigger <- function(x) {
  name <- x$NAME
  county_roads <- tigris::roads(state = x$STATEFP,
                                county = x$COUNTYFP, year = 2020)
  county_roads <- mutate(county_roads, NAME = name)
  
  return(county_roads)
}
```

```{r Pull in Road Data, warning = F, eval = F}
counties_filt <- tigris::counties(state = c('CO', 'UT'))
counties <- st_join(counties_filt, st_as_sf(FO), join = st_intersects, left = F)

# WE WILL NEED TO IMPORT JOHN_TIGGER FUNCTION FROM 'REVISIONARY TAXONOMY' - HOW YOU WANT DO THIS DORK? - IT COULD
# REQUIRE YOU ACTUALLY PUSHING A PACKAGE PUBLIC...

lst_result <- apply(counties, 1, john_tigger)
cnty_rds <- bind_rows(lst_result) %>% 
  st_as_sf()

FO_ch_15mi <- FO %>% 
  st_as_sf() %>% 
  st_transform(32613) %>% 
  st_buffer(16000.0934) %>% 
  st_transform(st_crs(cnty_rds)) %>% 
  select(2)

cnty_rds <- st_intersection(cnty_rds, FO_ch_15mi) # reduce extent.
inter <- st_intersects(cnty_rds,cnty_rds, sparse = TRUE)
cnty_rds <- cnty_rds[lengths(inter)>1,]

cnty_rds <- cnty_rds %>% 
  st_transform(32613) %>% 
  st_simplify(., preserveTopology = FALSE, dTolerance = 20)

ggplot(counties) +
  geom_sf() +
  geom_sf(data = st_as_sf(FO), fill = 'blue', alpha = 0.5) +
  geom_sf(data = cnty_rds, alpha = 0.7) +
  theme_bw()

st_write(cnty_rds, paste0(here(), '/data/processed/UFO_roads.shp'), quiet =T)
rm(lst_result, john_tigger, FO_ch_15mi, counties_filt, inter)
```


### Load 

### Locate most proximal road access to Points

Here we find the 5 segments of road which are most close to the point. This is to buffer against having to hike up and down a canyon.
```{r lcoate positions on most proximal stretches of road}
p1 <- paste0(here(), '/data/raw/', 
       list.files(paste0(here(), '/data/raw/'), pattern = 'PADUS'), "/")
PAD <- st_read(paste0(p1, list.files(p1, pattern = 'Fee.*\\.shp$')), quiet = T)
PAD_valid <- PAD %>% st_is_valid()
PAD[which(PAD_valid %in% F),] <- st_make_valid(PAD[which(PAD_valid %in% F),])

aim_pts<- st_read(paste0(here(), '/data/raw/AIM_Points_2022.shp'), quiet = T)
cnty_rds <- st_read(paste0(here(), '/data/processed/UFO_roads.shp'), quiet =T)
# dem <- terra::project(dem, "epsg:26912")
#writeRaster(dem, paste0(here(), '/data/processed/', "UFO_dem_10_smooth_UTM.tif"))
dem <- rast(paste0(here(), '/data/processed/', "UFO_dem_10_smooth_UTM.tif"))


road_parking <- perspective_routes(target_pts = aim_pts, roads = cnty_rds, dem = dem, 
                           search_dist = 4828, public_lands = PAD, road_segments = 400)


ggplot(road_parking) +
  geom_sf(aes(colour = location))

rm(perspective_routes, closest, nearestPt2points, p1, cnty_rds, FO, PAD_valid)
```

### Determine whether these roads follow public access or cross private without right of way


```{r gdistance example, eval = F}
r <- raster(system.file("external/maungawhau.grd", package="gdistance"))

altDiff <- function(x){x[2] - x[1]}
hd <- transition(r, altDiff, 16, symm=FALSE)
slope <- geoCorrection(hd)
adj <- adjacent(r, cells=1:ncell(r), pairs=TRUE, directions=16)
speed <- slope
speed[adj] <- 6 * exp(-3.5 * abs(slope[adj] + 0.05))
Conductance <- geoCorrection(speed) 

A <- c(2667670, 6479000)
B <- c(2667800, 6479400)
AtoB <- shortestPath(Conductance, A, B, output="SpatialLines")
BtoA <- shortestPath(Conductance, B, A, output="SpatialLines")

plot(r, xlab="x coordinate (m)", ylab="y coordinate (m)",
     legend.lab="Altitude (masl)")
lines(AtoB, col="red", lwd=2)
lines(BtoA, col="blue")
text(A[1] - 10, A[2] - 10, "A")
text(B[1] + 10, B[2] + 10, "B")

rm(altDiff, hd, slope, adk, speed, Conductance, A, B, AtoB, BtoA)
```



```{r worked examples}

# EXAMPLE PROCESS
ext1 <- road_parking %>% 
  filter(plot_id == 'AS-012') %>% 
  ext(.) %>% 
  as.polygons()              
empty1 <- terra::mask(dem, ext1, inverse = T)


# ACTUALY PROCESS
empty_raster <- rast(dem)
empty_raster <- crop(empty_raster, ext(crop_ext))
values(empty_raster) <- 1

areas <- road_parking %>% 
  st_buffer(250) %>% 
  split(., .$plot_id)
extents <- lapply((lapply(areas, ext)), as.polygons)

empty1 <- mapply(FUN = mask, empty, extents, inverse = T)

plot(empty1)


altDiff <- function(x){x[2] - x[1]}
hd <- transition(dem, altDiff, 16, symm=FALSE)
slope <- geoCorrection(hd)
adj <- adjacent(dem, cells=1:ncell(r), pairs=TRUE, directions=16)
speed <- slope
speed[adj] <- 6 * exp(-3.5 * abs(slope[adj] + 0.05))
Conductance <- geoCorrection(speed) 

parking <- park2point %>% 
  filter(location == 'parking')
point <- park2point %>% 
  filter(location == 'point') 

to_plot <- shortestPath(Conductance, parking, point, output="SpatialLines")
from_plot <- shortestPath(Conductance, point, parking, output="SpatialLines")

```