---
title: "ESD_climate_physiographic_metrics"
author: "steppe"
date: '2022-05-11'
output:
  pdf_document: default
  html_document: default
---

An ecological site description is often based upon several abiotic parameters. The most common of these are:

Physiographic:
  1) Elevation
  2) Aspect
  3) Slope

Climate:
  1) Annual Air Temperature
  2) Annual Total Precipitation 
    a) Timing of Precipitation
    b) Driest & Wettest Months
  3) Frost free period
  4) Initiation and cessation of vegetative growth


```{r Load Libraries, echo = F, warning = F, message = F}
library(tidyverse)
library(sf)
library(terra)
library(prism)
library(here)
```

# Data Analysis

```{r import aim pts, echo = F}
aim_pts <- st_read(paste0(here(), '/data/raw/AIM_Points_2022.shp'), quiet = T) %>% 
  st_transform(26912) %>% 
  st_buffer(30) %>% 
  st_transform(4269)
```

## Physiographic
```{r Extract physiographic variables, echo = F}
slope <- rast(paste0(here(), '/data/processed/',
                     list.files(paste0(here(), '/data/processed'),
                                pattern = 'slope.*\\.tif')
                     )
              )

aspect <- rast(paste0(here(), '/data/processed/',
                      list.files(paste0(here(), '/data/processed'),
                                 pattern = 'aspect.*\\.tif')
                      )
               )

dem <- rast(paste0(here(), '/data/processed/',
                   list.files(paste0(here(), '/data/processed'), 
                              pattern = "dem.*smooth*\\.tif") 
                   )
            )

aim_pts_v <- vect(aim_pts)

physio <- tibble(
  slope = extract(slope, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2],
  dem = extract(dem, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2],
  aspect = extract(aspect, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2]
) %>% 
  mutate('elev_ft' = dem * 3.28084) %>% 
  mutate('slope_prcnt' = tan(slope*pi/180)*100) %>% 
  dplyr::select(-slope, -dem)

rm(dem, aspect, slope)
```

## Climatic

```{r Download relevant data from PRISM for Annual Metrics, echo = F, message = F, warning = F}
#dir.create(paste0(here(), '/data/raw/prism'))
prism_set_dl_dir(paste0(here(), '/data/raw/prism'))

get_prism_normals("ppt", "800m", annual = TRUE, keepZip = FALSE)
get_prism_normals("tmean", "800m", annual = TRUE, keepZip = FALSE)

path <- paste0(here(), '/data/raw/prism/')

tmean <- rast(paste0(paste0(path, list.files(path, pattern = 'tmean'), '/'), 
       list.files(paste0(path, list.files(path, pattern = 'tmean'), '/'), 
                  pattern = '.bil$')))
ppt <- rast(paste0(paste0(path, list.files(path, pattern = 'ppt.*annual'), '/'), 
       list.files(paste0(path, list.files(path, pattern = 'ppt.*annual'), '/'), 
                  pattern = '.bil$')))

annual_climate <- tibble(
  tmean_C = extract(tmean, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2],
  ppt_mm = extract(ppt, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2],
) %>% 
  mutate(ppt_inch = (ppt_mm / 10) * 0.393701) %>% 
  mutate(tmean_F = (tmean_C * 9/5) + 32) %>% 
  dplyr::select(-ppt_mm)
  

rm(tmean, ppt)
```

```{r Download relevant data from PRISM for Monthly Metrics, echo = F, message = F}

get_prism_normals(type = "ppt", resolution = "800m", mon = 1:12, keepZip = FALSE)

ppt <- rast(paste0(paste0(path, list.files(path, pattern = 'ppt.*M3_[0-9]{2}'), '/'), 
       list.files(paste0(path, list.files(path, pattern = 'ppt.*M3_[0-9]{2}'), '/'), 
                  pattern = '.bil$')))
names(ppt) <- substr(month.name, 1, 3) 

monthly_precip <- extract(ppt, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2:13] 
monthly_precip <- data.frame(monthly_precip/10)

monthly_precip <- monthly_precip %>% 
  mutate('annual_total' = rowSums(.)) %>% 
  mutate(across(.cols = 1:12, ~ round((.x/annual_total)*100, 1)))

rm(ppt)
```

```{r frost free period, echo = F}

frost_free <- rast(paste0(here(), '/data/raw/FrostFreeDays/',
                      list.files(paste0(here(), '/data/raw/FrostFreeDays'),
                                 pattern = 'historical.*\\.tif')
                      )
               )

ff <- tibble(Frost_free = extract(frost_free, aim_pts_v, method = 'bilinear', fun = mean, weights = T)[,2])
rm(frost_free)
```


# Results

```{r, echo = F, message = F, warning = F}
aim_pts <- aim_pts %>% 
  st_drop_geometry() %>% 
  select(plot_id)
  
cbind(aim_pts, physio, annual_climate, ff) %>% 
  mutate(across(where(is.numeric), ~ round(.,0))) %>% 
  dplyr::select(plot_id, elev_ft, slope_prcnt, aspect, ppt_inch, tmean_C, tmean_F, Frost_free) %>% 
  kableExtra::kable(., "latex", booktabs = TRUE, longtable = TRUE, 
                    caption = "Some AIM site characteristics 2022 UFO",
                    col.names = c('Plot ID', 'Elev. (ft.)', '% Slope', 'Aspect', 'Ppt. (in.)', 'Annual Temp. (C)',  
                                  'Ann. Temp. (F)', 'Frost Free (d)')) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position", "repeat_header"))
```

