---
title: "Gather ESD's for AIM"
author: "steppe, Michael Schmidt, K. Holsinger"
date: '2022-04-25'
output:
  pdf_document: default
  html_document: default
---

```{r, echo = F}
knitr::opts_chunk$set(
  echo = FALSE,  
  message = FALSE, 
  warning = FALSE 
)
```

```{r Load Libraries, echo = F, warning = F, message = F}
library(tidyverse)
library(here)
library(janitor)
library(sf)
source(here::here('scripts/functions.R'))
```


Ecological Site Descriptions serve to identify repeating motif across the landscape which are ecologically similar and respond to management in parallel fashions. 

Traditionally, ESD's may be downloaded manually from the Jornada website via an NRCS soil mapping website which links to them. However this process can become tedious. Here I show a method for organization to generate their own script for:

1) Downloading and setting up the MLRA query box, you can also do this via 
2) Downloading all ESD's from the Jornada
3) Generating a list of ESD'S within a certain distance of a location (e.g. AIM plot)

# Download MLRA Database if your are unsure which region your field office is in

[NRCS](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_053624)

I moved this file to my data/raw sub-directory, and renamed it 'nrcs_mlra' because it's initial name did not indicate it's function.

```{r Import Starting Data}
blm_admnu <- st_read(paste0(here(), '/data/raw/blm_co_admu_fo_shp/',
  list.files(paste0(here(), '/data/raw/blm_co_admu_fo_shp'), pattern = '.shp$')), 
  quiet = T)
here()

mlra <- st_read(paste0(here(), '/data/raw/nrcs_mlra/',
  list.files(paste0(here(), '/data/raw/nrcs_mlra'), pattern = '.shp$')), 
  quiet = T)

head(blm_admnu[,c(2,7)]) %>% 
  st_drop_geometry()
```

I work in the Uncompahgre field office, hence I will filter for that administrative name ('ADMU_NAME'). An alternative it to filter by the District office ('PARENT_NAME'), for example the 'Southwest District Office'; note both of these are in all caps. 

```{r Filter Administrative Unit or Parent Name to relevant area}
blm_admnu <- blm_admnu  %>% 
  filter(ADMU_NAME == 'UNCOMPAHGRE FIELD OFFICE') %>% 
  st_transform(32613) %>%  # we transform to a PLANAR projection in meters so we can buffer it
  st_buffer(10000) %>%  # buffer the FO by 10 km  
  # this allows us to retrieve bordering MLRAS
  st_transform(st_crs(mlra)) # convert new vector data to same projection as mlra

mlra_lookup <- st_intersects(blm_admnu, mlra, sparse = T)
mlra <- mlra[mlra_lookup[[1]],]

ggplot(mlra) +
  geom_sf(aes(colour = MLRA_NAME)) +
  geom_sf(data = blm_admnu, fill = 'yellow', alpha = 0.4) +
  theme_bw()

rm(mlra_lookup)
```

## Create list of all ESD's for MLRA's

```{r Create list of ESDs, eval = F}
mlra$MLRARSYM <- paste0('0', mlra$MLRARSYM) # these are meant to be preceded by a leading 0
mlra$MLRARSYM[3] <- paste0(mlra$MLRARSYM[3], 'X') # 36 is suppose to have a trailing X.... this only applies to me!!!!!

list_url <- paste0("https://edit.jornada.nmsu.edu/services/downloads/esd/", mlra$MLRARSYM, "/class-list.txt")
```

# Download all Relevant ESD from Jornada

Using some code fleshed out by Michael Schmidt we can download the files below. 
```{r Download ESDs, eval = F}
ecoclasses <- do.call(rbind,
          lapply(list_url, read_tsv, skip = 2, show_col_types = FALSE)) %>% 
  janitor::clean_names()

# NOW FILTER TO YOUR STATE OF INTEREST, for example our MLRA spans several States
co_classes<-ecoclasses %>%
  filter(str_detect(ecological_site_id, "CO|UT$"))

download_esd<-function(x, output_path){
  initial_timeout <- getOption('timeout')
  options(timeout=7200)
  base_url <- "https://edit.jornada.nmsu.edu/services/descriptions/esd/"
  doc_url <- paste0(base_url, x$mlra, "/", x$ecological_site_id, ".pdf")
  download.file(doc_url, paste0(output_path,"/",x$ecological_site_id," - ",
                                janitor::make_clean_names(x$ecological_site_name),".pdf"), mode="wb")
  Sys.sleep(3)
  options(timeout = initial_timeout)
}

dirpath <- paste0(here(), '/data/raw/UFO_ESD') # MODIFY THIS TO be the folder you want to hold your esd
# dir.create(dirpath) # this creates a folder to hold the ESD 

download_esd(co_classes, dirpath)
rm(mlra, ecoclasses, co_classes, dirpath, list_url, download_esd)
```

# 3 Generate a list of the relevant ESD's for each AIM point

We will just write these out as an excel like file. 

You can hopefully  download the state SSURGO database from here: 
[NRCS SSURGO BOX](https://nrcs.app.box.com/v/soils/folder/148414960239)

Be sure to check if your state has the 'spatial' component with it- if not it is a not go. For Colorado I had to individually download counties/areas and generate a reference data set from [here](https://websoilsurvey.sc.egov.usda.gov/App/WebSoilSurvey.aspx). 

```{r Link to SSURGO database, eval = F}
sf::st_layers(dsn = paste0(here(), '/data/raw/gSSURGO_CO/gSSURGO_CO.gdb'))

st_read(dsn = paste0(here(), '/data/raw/gSSURGO_CO/gSSURGO_CO.gdb'), layer = 'mapunit') # ESDs live here
table(db$siteindex_l)
```

For the UFO we had to create our own lookup data file...
```{r Generate novel ssurgo dataset, eval = F}

spatial <- paste0(here(), '/data/raw/ssurgo/',
                  list.files(paste0(here(), '/data/raw/ssurgo/'), 
                             pattern  = 'soilmu_a.*.shp$', recursive = T))
component <- paste0(here(), '/data/raw/ssurgo/',
                  list.files(paste0(here(), '/data/raw/ssurgo/'), 
                             pattern = 'comp', recursive = T))
ecoclass <- paste0(here(), '/data/raw/ssurgo/',
                  list.files(paste0(here(), '/data/raw/ssurgo/'), 
                             pattern = 'cecoclas.txt', recursive = T))

spatial <- do.call(bind_rows,
        lapply(spatial, st_read, quiet = T))
component <- do.call(rbind,
          lapply(component, read_delim, delim = "|", col_names = F, show_col_types = FALSE))[,108:109]
names(component) <- c('mukey','cokey')
ecoclass <- do.call(rbind,
          lapply(ecoclass, read_delim, delim = "|", col_names = F, show_col_types = FALSE))
names(ecoclass) <-  c('ecoclasstypename', 'ecoclassref', 'ecoclassid', 'ecoclassname', 'cokey', 'coecoclasskey')

ecoclass <- left_join(ecoclass, component, by = 'cokey') %>% 
  mutate(mukey = as.character(mukey))

ssurgo <- left_join(spatial, ecoclass, by = c('MUKEY' = 'mukey')) 

blm_admnu <- st_transform(blm_admnu, st_crs(ssurgo))

ssurgo_valid <- st_is_valid(ssurgo)
invalid <- as.numeric(which(ssurgo_valid %in% F))
ssurgo[invalid,] <- st_make_valid(ssurgo[invalid,])

ssurgo_lookup <- st_intersects(blm_admnu, ssurgo, sparse = T)
ssurgo <- ssurgo[ssurgo_lookup[[1]],] %>% 
  select(7:11)

st_write(ssurgo, paste0(here(), '/data/processed/ssurgo.shp'), quiet = T)
rm(ecoclass, component, ssurgo_valid, ssurgo_lookup, invalid, spatial)
```


```{r Extract soil survey areas to points, eval = F}

ssurgo <- st_read(paste0(here(), '/data/processed/ssurgo.shp'), quiet = T)
aim_plts_buff<- st_read(paste0(here(), '/data/raw/AIM_Points_2022.shp'), quiet = T) %>% 
  st_transform(26912) %>% 
  st_buffer(1610) %>% 
  st_transform(st_crs(ssurgo))

esds <- st_intersection(ssurgo, aim_plts_buff) %>% 
  group_by(plot_id, ecclssd, ecclssn) %>% 
  summarize(geometry = st_union(geometry)) 

ggplot(esds, aes(fill = ecclssd), color = NA) +
  geom_sf() 
 
esds$Area <- round( 
    (units::set_units(st_area(esds), km^2) / # caluclate area of each mapped ESD
       mean(units::set_units(st_area(aim_plts_buff), km^2)) ) # standardize areas to mean area of buffer
    * 100, # convert to percentile
    1) 

dir.create(paste0(here(), '/data/processed/ESDS_spatial'))
st_write(esds, paste0(here(), '/data/processed/ESDS_spatial/esds.shp'), quiet = T)
```

```{r Write table of ESD by plot}

esds <- st_read(paste0(here(), '/data/processed/ESDS_spatial/esds.shp'), quiet = T)

esds %>% 
  group_by(plot_id) %>% 
  arrange(plot_id, -Area) %>% 
  mutate(ecclssn = str_to_title(ecclssn)) %>% 
  collapse_rows_df(plot_id) %>% 
  st_drop_geometry() %>% 
  kableExtra::kable(., "latex", booktabs = TRUE, longtable = TRUE, 
                    caption = "Ecological Sites Descriptions 2022 UFO AIM", 
                    col.names = c('Plot ID', 'Eco Site Code', 'Eco Site Name', 'Percent Area')) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header")) 

```

