---
title: "Sample Points & Office Rejections"
author: "steppe"
date: '2022-04-27'
output: html_document
---

The purpose of this script is to assist a crew lead in searching for office rejections in the pre-season. It will use a simple set of common spatial data products (e.g. a Digital Elevation Model, BLM Land Ownership, and Roads) to flag points for an office rejection. It is my (many to most persons) firm belief that all points should be investigated regardless after running this script using Google Earth or other 3D rendering engine. We hope that this script:

1) Reduces the time required to import a KML dataset (of AIM points by statum and position in the sample design) to Google Earth.
2) Allows you to readily create and import BLM ownership vector data as KML to Google Earth.
3) Reduces uncertainty in confirming office rejections with the computed metrics which are then verified in Google Earth by you - the analyst. 

# Background Data Handling

```{r Load Libraries, echo = F, warning = F, message = F}
library(tidyverse)
library(readxl)
library(here)
library(sf)
library(terra)
library(leaflet)
library(maptools)
source(here::here('scripts/functions.R'))
```

### Create a copy of your tracking sheet. 

My advice is to to create a MASTER version of all points, and then copy this to a COPY of all points
```{r Copy AIM Sample Design Sheet}

Plot_Tracking_2022 <- read_excel(paste0(here(),  '/data/raw/PlotTracking_AIM_plots_2022-COPY.xlsx'), 
                           sheet = 1, skip = 1) %>% 
  janitor::clean_names()

Proportions <- read_excel(paste0(here(),  '/data/raw/PlotTracking_AIM_plots_2022-COPY.xlsx'), 
                           sheet = 2, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(design_strata, plots)

initial_process <- read_excel(paste0(here(),  '/data/raw/PlotTracking_AIM_plots_2022-COPY.xlsx'), 
                           sheet = 3) %>% 
  janitor::clean_names() 

initial_process <- initial_process %>% 
  drop_na(long:lat) %>% 
  filter(long != 'SAMPLED IN 2018') %>% 
  st_as_sf(coords = c(x = 'long', y = 'lat'), crs = 4269, remove = F) 
```

### Slope Rejections

AIM points are to be rejected if they fall upon a slope of 50% or greater. A 50% slope is 26.57 degrees. This value is somewhat notoriously hard to calculate with a DEM, but we had near perfect concordance between our method and an expert analyst (actually we caught 2 more missed rejections via this).

```{r Office Reject Slopes which are too great to Sample}

slope_degree <- rast(paste0(here(), '/data/processed/UFO_slope_deg.tif'))
dem <- raster(paste0(here(), '/data/processed/UFO_dem_10_smooth.tif'))

ouput <- plot_mover(initial_process, utm_nad = 26913,
                    ras_dem = dem, ras_slo = slope_degree)

ouput <- ouput %>% 
  mutate(across(.cols = long:lat, as.numeric))
```


### Elevation of points 

The lower the land the hotter the site. Sample these and work up.
```{r}

outro <- ouput %>% 
  mutate(Elevation = extract(dem, .))

sort(outro)
```


```{r}
twop <- ouput %>% 
  filter(next_in_line_or_oversample == 'Yes',
         str_detect(plot_id, 'SS|SD'))

leaflet(data = twop) %>%
 addTiles() %>%
 addMarkers(~long, ~lat,
            popup = ~as.character(plot_id),
            label = ~as.character(plot_id))
```

# Prepare Data for Entry to Google Earth

Google Earth provides an excellent interface to visualize sites due to it's 3-dimensional interface. We will prep and style the points as kml formats for quick input to the system. 

### Create Site Icons for Google Earth

```{r Save AIM Points as Icons}
base_url <- 'http://google.com/mapfiles/ms/micons/'
url_suffix <- '.png'

opts_base <- paste0(base_url, c('yellow', 'blue', 'green', 'lightblue', 
               'orange', 'pink', 'purple', 'red'), url_suffix)
opts_oversamples <- paste0(base_url, 
                           c('yellow-dot', 'blue-dot', 'green-dot', 'ltblue-dot', 
                             'orange-dot', 'pink-dot', 'purple-dot', 'red-dot'),
                           url_suffix)

rm(base_url, url_suffix)
```



### Create Ownership Layers for Google Earth

Here we will use the PAD-US database to download public land ownership data by state. Note it is almost always not imperative that you check that your points fall on BLM (the chance of this is **very very low**), what matters is crossing private land without easements to access BLM land. 

You can download the [PAD-US dataset](https://www.sciencebase.gov/catalog/item/60259839d34eb12031138e1e) for your state manually. 

```{r Import Public Lands owernship and crop to extent of Field Office. echo = F}

p1 <- paste0(here(), '/data/raw/',
  list.files(paste0(here(), '/data/raw/'), pattern = '^PADUS'), '/'
  ) # list folder for state
PAD <- st_read(paste0(p1, list.files(p1, pattern = 'Fee.*\\.shp$')), quiet = T)
PAD_valid <- PAD %>% st_is_valid()

PAD[which(PAD_valid %in% F),] <- st_make_valid(PAD[which(PAD_valid %in% F),])

blm_admnu <- st_read(paste0(here(), '/data/raw/blm_co_admu_fo_shp/',
  list.files(paste0(here(), '/data/raw/blm_co_admu_fo_shp'), pattern = '.shp$')), 
  quiet = T) %>% 
  filter(ADMU_NAME == 'UNCOMPAHGRE FIELD OFFICE') %>% 
  st_transform(32613) %>%  # we transform to a PLANAR projection in meters so we can buffer it
  st_buffer(10000) %>% 
  st_transform(st_crs(PAD)) %>% 
  st_bbox()

PAD <- st_crop(PAD, blm_admnu)

head(PAD)
table(PAD$Mang_Type) # RELEVANT
table(PAD$Mang_Name)
table(PAD$Des_Tp)

PAD <- PAD %>% 
  dplyr::select(Mang_Type, Mang_Name, Des_Tp) %>% 
  mutate(Mang_Name = case_when(
    Mang_Name %in% c('CITY', 'CNTY', 'SDNR', 'SFW', 'SPR', 'OTHS', 'REG') ~ 'CITY_CNTY_SDC_SDNR_SPR', 
    Mang_Name %in% c('NGO','PVT') ~ 'PVT',
    TRUE ~ Mang_Name
  ))

PAD1 <- PAD %>% 
  group_by(Mang_Name) %>% 
  summarize(geometry = st_union(geometry))

ggplot(PAD1, aes(fill = Mang_Name)) +
  geom_sf()
  
rm(p1, PAD_valid)
```

I find that these are the relevant tables for this process, you may locate them in their original form on the
[PAD-US Data Manual](https://www.usgs.gov/programs/gap-analysis-project/pad-us-data-manual) webpage. Note that I removed many island and oceanic entities from Table 5. 

Table 4. Agency Type domain codes and descriptions

| Code |	      Domain Description         |
| ---- | --------------------------------- |
| DESG |	Designation                      | 
| DIST |	Regional Agency Special District |
| FED  |  Federal                          |
| JNT  |	Joint                            |
| LOC  |	Local Government                 |
| NGO  |	Non-Governmental Organization    |
| PVT  |	Private                          |
| STAT |  State                            |
| TERR |	Territorial                      |
| TRIB |	American Indian Lands            |
| UNK  |	Unknown                          |

Table 5. Agency Name domain codes and descriptions

| Code  |            Domain Description             |
| ----  | ----------------------------------------- |
| BIA   | Bureau of Indian Affairs                  |
| BLM   | Bureau of Land Management                 |
| CITY  | City Land                                 |
| CNTY 	| County Land                               |
| DESG 	| Designation                               |
| DOD 	| Department of Defense                     |
| DOE 	| Department of Energy                      |
| FWS 	| U.S. Fish & Wildlife Service              |
| JNT 	| Joint                                     |   
| NGO 	| Non-Governmental Organization             |
| NPS 	| National Park Service                     |
| NRCS  | Natural Resources Conservation Service    |
| OTHF 	| Other or Unknown Federal Land             |
| OTHR 	| Other                                     |
| OTHS 	| Other or Unknown State Land               |
| PVT 	| Private                                   |
| REG 	| Regional Agency Land                      |
| RWD 	| Regional Water Districts                  |
| SDC 	| State Department of Conservation          |
| SDNR 	| State Department of Natural Resources     |
| SDOL 	| State Department of Land                  |
| SFW 	| State Fish and Wildlife                   |
| SLB 	| State Land Board                          |
| SPR 	| State Park and Recreation                 |
| TRIB 	| American Indian Lands                     |
| UNK 	| Unknown                                   |
| UNKL 	| Other or Unknown Local Government         |
| USACE | Army Corps of Engineers                   |
| USBR 	| Bureau of Reclamation                     |
| USFS 	| Forest Service                            |



```{r Save Public Lands Ownership as KML}

kmlfile <- paste(td, "BLM_UFO_Ownership.kml", sep="/")

public_lands_colours <- setNames(
  
  # these manually transcribed from the H-1553-Publications Standards Manual
  # Handbook - hopefully no errors.
  # [H-1553](https://www.ntc.blm.gov/krc/uploads/223/Ownership_Map_Color_Reference_Sheet.pdf)
  
  c( # colours
  rgb(254, 230, 121, max = 255), # BLM
  rgb(204, 235, 197, max = 255), # USFS
  rgb(202, 189, 220, max = 255), # NPS
  rgb(127, 204, 167, max = 255), # FWS
  rgb(255, 255, 179, max = 255), # USBR
  rgb(253, 180, 108, max = 255), # TRIB
  rgb(251, 180, 206, max = 255), # DOD
  rgb(228, 196, 159, max = 255), # OTHF
  rgb(179, 227, 238, max = 255), # SLB
  rgb(255, 255, 255, max = 255), # PVT
  rgb(143, 181, 190, max = 255) # CITY CNTY
  ), 
  
  c( # names
  'BLM', 'USFS', 'NPS', 'FWS', 'USBR', 'TRIB', 'DOD', 'OTHF', 'SLB', 'PVT', 'CITY_CNTY_SDC_SDNR_SPR'
  )
)

public_lands_colours

kmlPolygons(blm_ownership, kmlfile = kmlfile,
  name = "KML Polygons subset", 
  description = "three countries", 
  col = "blue",
  visibility = 1, lwd = 1, border = "white", kmlname = "R Test 2",
  kmldescription = "This is <b>only</b> a <a href='http://www.r-project.org'>R</a> test.")
## combine to make a list of polygon objects to plot
polList <- c(regions,wrld_simpl)
kmlfile <- paste(td, "worldPoliticalandSubset.kml", sep="/")
```



U.S. Geological Survey (USGS) Gap Analysis Project (GAP), 2020, Protected Areas Database of the United States (PAD-US) 2.1: U.S. Geological Survey data release, https://doi.org/10.5066/P92QM3NT.
